---
title: "maps-bovid-ecometrics"
author: "Andrew Barr"
date: "September 15, 2015"
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
---

## Preliminaries

```{r warning=FALSE, message=FALSE}
library(rgdal)
library(rgeos)
library(maptools)
library(dplyr)
library(raster)
library(tidyr)
library(reshape2)
library(GISTools)

par.original <- par(no.readonly=TRUE)

summarize_regression <- function(model, round=2, label=""){
  mod_summary <- summary(model)
  return(knitr::kable(data.frame(
    label=label,
    intercept=round(mod_summary$coefficients[1,1],round),
    SEI=round(mod_summary$coefficients[1,2],round),
    slope=round(mod_summary$coefficients[2,1],round),
    SES=round(mod_summary$coefficients[2,2],round),
    r2 = round(mod_summary$adj.r.squared,round)
    )))
}

summarize_ANOVA <- function(model, round=2, label=""){
  mod_summary <- summary(model)
  return(knitr::kable(data.frame(
    label=label, 
    ## a litte tricky to get correct coeffs
    # because the coefficients are in the original order, even though I used a ordered factor in the lm()
    forest=round(mod_summary$coefficients[1,1],round),
    wetland=round(mod_summary$coefficients[6,1],round),
    woodySav=round(mod_summary$coefficients[3,1],round),
    sav=round(mod_summary$coefficients[4,1],round),
    openShrub=round(mod_summary$coefficients[2,1],round),
    grass=round(mod_summary$coefficients[5,1],round),
    barren=round(mod_summary$coefficients[7,1],round),
    r2 = round(mod_summary$adj.r.squared,round)
    )))
}
```

## Load africa

```{r}
#get base file path to load data files...assumes file paths are from perspective of directory containing this file
base_path <- getwd()
africa <- readOGR(paste0(base_path,"/africa_shp/africa.shp"), "africa")
```

## Load bovids and merge based on taxonomic name

```{r cache=FALSE}
bovids <- readOGR(paste0(base_path,"/IUCN-Redlist/AFRICAN_BOVIDS.shp"), "AFRICAN_BOVIDS")
bovids <- spTransform(bovids, proj4string(africa))

mergedBovidPolys <- unionSpatialPolygons(bovids, IDs = bovids@data$id_no)
mergedBovidAttributes <- group_by(bovids@data, id_no) %>% summarize(binomial = unique(binomial))
class(mergedBovidAttributes) <- "data.frame"
rownames(mergedBovidAttributes) <- mergedBovidAttributes$id_no
bovids <- SpatialPolygonsDataFrame(mergedBovidPolys, mergedBovidAttributes)
rm(mergedBovidAttributes,mergedBovidPolys)

#eliminate outlier taxa, because they are is usually the only taxon in the community and are north of Sahara
##
outliers <- c(which(bovids@data$binomial == "Ammotragus lervia"), which(bovids@data$binomial == "Gazella cuvieri"), which(bovids@data$binomial == "Addax nasomaculatus"))
bovids <- bovids[-outliers,]

```



## Make raster 

```{r cache=FALSE}
africaRaster <- raster(extent(africa), res=c(50000,50000), crs=CRS(proj4string(africa)))
africaRaster <- mask(africaRaster, africa)
```

## Compute bovid species richness

```{r cache=FALSE}
bovidRichness <- rasterize(x=bovids, y=africaRaster, fun='count', field="id_no")
plot(bovidRichness, axes=F, box=F, main="Bovid Species Richness")
```


## Get data from sqlite ecomorph database, clone of production ecomorph database on webfaction

```{r cache=FALSE}
joinedDFs <- read.table("all_ecomorphic_data.csv", sep=",", header=T)
bovids@data$B <- joinedDFs$B/joinedDFs$geomeans
bovids@data$DistRad <- joinedDFs$DistRad/joinedDFs$geomeans
bovids@data$ProxRad <- joinedDFs$ProxRad/joinedDFs$geomeans
bovids@data$MML <- joinedDFs$MML/joinedDFs$geomeans
bovids@data$WAF <- joinedDFs$WAF/joinedDFs$geomeans

bovids@data$relMidshaftWidth <- joinedDFs$relMidshaftWidth
bovids@data$relMetarsalLength <- joinedDFs$relMetarsalLength
bovids@data$relMetatarsalDepth <- joinedDFs$relMetatarsalDepth
bovids@data$medialVert <- joinedDFs$medialVert
bovids@data$distalArticular <- joinedDFs$distalArticular
```


## Load MAP raster

```{r}
MAP <- raster(paste0(base_path,"/globalMAP_2point5_arcseconds.grd"))
MAP <- projectRaster(MAP, africaRaster)
```



## Load land cover raster

```{r}
landCover <- raster(paste0(base_path,"/AFRICA_LandCover.TIFF"))
landCover <- projectRaster(landCover, africaRaster, method="ngb")
values(landCover)[values(landCover) %in% c(0, 1, 3, 4, 5, 6, 12, 13, 14)] <- NA
#method = 'ngb' refers to nearest neighbor, useful for categorical rasters
```


## function to do a single variable

```{r}
doEcometricAnalysis <- function(var, 
                                originalMAP, 
                                originalLandCover, 
                                savePlots=FALSE, 
                                dir=base_path, 
                                width=5, height=5, 
                                cex.main=1.75, 
                                plotLabel=NULL, 
                                scaleBar=FALSE,
                                northArrow=FALSE,
                                returnLMs=TRUE,
                                writeEcometricRaster=FALSE) {
  stopifnot(var %in% names(bovids@data))
  ecometric <- rasterize(x=bovids, 
                         y=africaRaster, 
                         fun=function(x, ...) {mean(x, na.rm=TRUE)}, 
                         field=var)
  plot(ecometric,axes=F,box=F, main=paste("mean", var))
  if(writeEcometricRaster) writeRaster(ecometric, paste0(base_path,"/ecometricRasters/",var,".tif"), overwrite=TRUE)

  MAP <- mask(originalMAP, ecometric)
  
  MAPlm <- lm(scale(values(ecometric))~sqrt(values(MAP)))
  summary(MAPlm)
  plot(scale(values(ecometric))~sqrt(values(MAP)), pch=16, xlab="sqrt(MAP)", ylab=var)
  abline(reg=MAPlm, col="red", lty=2)
  
  landCover <- mask(originalLandCover, ecometric)
  landCoverFactor <- factor(values(landCover))
  levels(landCoverFactor) <- c("Evergreen\nBroad. Forest",
                               "Open\nShrub.",
                               "Woody\nSavannas",
                               "Savannas",
                               "Grasslands",
                               "Perm.\nWetland",
                               "Barren")
  landCoverFactor <- ordered(landCoverFactor, 
                     levels=c("Evergreen\nBroad. Forest", "Perm.\nWetland","Woody\nSavannas", "Savannas","Open\nShrub.","Grasslands",  "Barren"))
  plot(scale(values(ecometric))~landCoverFactor, col=(terrain.colors(7)), las=2, xlab="", ylab="", main=var)
  box()

    
  if(savePlots){
    pdf(file = paste0(dir,"/",var,".pdf"), width=width, height=height)
    plot(scale(ecometric), axes=FALSE, box=F)
    text(1636265, 3249920, labels=ifelse(is.null(plotLabel), var, plotLabel), cex=cex.main)
    if(scaleBar){      
      scalebar(c(-1795336,-911192.4), d=1000000, divs=2, label="1000 km", adj=c(0,-1),type="bar")
      north.arrow(-1198181, -1670000,len = 100000, lab = "N")
    }
    dev.off()
    
    pdf(file = paste0(dir,"/",var,"-MAPscatterplot.pdf"), width=width, height=height)
    plot(scale(values(ecometric))~sqrt(values(MAP)), pch=16, xlab="sqrt(MAP)", ylab=ifelse(is.null(plotLabel), var, plotLabel), main=ifelse(is.null(plotLabel), var, plotLabel), cex.main=cex.main)
    abline(reg=MAPlm, col="red", lty=5, lwd=3)
    dev.off()
    
    pdf(file = paste0(dir,"/",var,"-landcoverBoxplots.pdf"), width=width*1.8, height=height)
    plot(scale(values(ecometric))~landCoverFactor, col=(terrain.colors(7)), xlab="", ylab="", main=ifelse(is.null(plotLabel), var, plotLabel), cex.main=cex.main, cex.lab=0.5)
    dev.off()

    }
  
  print(summarize_regression(MAPlm, 2, label=sprintf("MAP~%s", var)))
  
  options(contrasts = c("contr.treatment", "contr.treatment")) #nesseary to have treament contrasts for ordered factors
  LClm <- lm(scale(values(ecometric))~landCoverFactor)
  print(summarize_ANOVA(LClm, 2, label=sprintf("LC~%s", var)))
  if(returnLMs) return(list(unscaledLC=lm(values(ecometric)~landCoverFactor), 
                            unscaledMAP=lm(values(ecometric)~sqrt(values(MAP)))
                            )
                      )
  }
```


## B

```{r results="asis", cache=TRUE}
resB <- doEcometricAnalysis("B", MAP, landCover, savePlots = F)
resB
```

## DistRad

```{r results="asis", cache=TRUE}
resDistRad <- doEcometricAnalysis("DistRad", MAP, landCover, savePlots = F)
resDistRad
```

## ProxRad

```{r results="asis", cache=TRUE}
resProxRad <- doEcometricAnalysis("ProxRad", MAP, landCover, savePlots = F)
resProxRad
```

## WAF

```{r results="asis", cache=TRUE}
resWAF <- doEcometricAnalysis("WAF", MAP, landCover, savePlots = F)
resWAF
```

## MML

```{r results="asis", cache=TRUE}
resMML <- doEcometricAnalysis("MML", MAP, landCover, savePlots = F, scaleBar=TRUE, northArrow=TRUE, plotLabel="MMLa")
resMML
```

## relMidshaftWidth

```{r results="asis", cache=TRUE}
resrelMidshaftWidth <- doEcometricAnalysis("relMidshaftWidth", MAP, landCover, savePlots = F, plotLabel="MMLm")
resrelMidshaftWidth
```

## relMetarsalLength

```{r results="asis", cache=TRUE}
resrelMetarsalLength <- doEcometricAnalysis("relMetarsalLength", MAP, landCover, savePlots = F, plotLabel="MLEN")
resrelMetarsalLength
```

## relMetatarsalDepth

```{r results="asis", cache=TRUE}
resrelMetatarsalDepth <- doEcometricAnalysis("relMetatarsalDepth", MAP, landCover, savePlots = F, plotLabel="MDepth")
resrelMetatarsalDepth
```

## medialVert

```{r results="asis", cache=TRUE}
resmedialVert <- doEcometricAnalysis("medialVert", MAP, landCover, savePlots = F, plotLabel="MVAP")
resmedialVert
```

## distalArticular

```{r results="asis", cache=TRUE}
resdistalArticular <- doEcometricAnalysis("distalArticular", MAP, landCover, savePlots = F, plotLabel="IDML")
resdistalArticular
```
